Soss

### 2019-12-29 paper/paper.tex
edits

\begin{abstract}
Soss is a probabilistic programming language embedded probabilistic programming language embedded in the Julia language. Soss represents statistical models in terms of abstract syntax trees, and uses staged compilation for on-demand generation of ``inference primitives'' (random sampling, log-density, etc) without reqquiring casual users to worry about such details.
\end{abstract}

@@

\begin{lstlisting}[language = Julia]
  using Plots

  x = -3.0:0.01:3.0
  y = rand(length(x))
  plot(x, y)
\end{lstlisting}

@@

\section{Symbolic Simplifcation}

\section{Code Generation}

>>>>

\begin{abstract}
We present Soss, a probabilistic programming language embedded in the Julia language. Soss represents statistical models in terms of abstract syntax trees, and uses staged compilation for on-demand generation of ``inference primitives'' (random sampling, log-density, etc) without reqquiring casual users to worry about such details.

The approach taken by Soss makes it easy to extend to take advantage of other packages in the rapidly-growing Julia ecosystem. At the time of this writing, Soss users can choose from several inference back-ends and connect easily with larger systems SymPy and Gen.
\end{abstract}

@@

\begin{lstlisting}[language = Julia]
m = @model n begin
    x ~ Normal(0,10) |> iid(n)
    α ~ Normal()
    β ~ Normal()
    σ ~ HalfNormal()
    yhat = α .+ β .* x
    y ~ For(n) do j 
            Normal(yhat[j],σ)
        end
end
\end{lstlisting}

@@

\section{Symbolic Manipulation}

\subsection{Symbolic Log-density}

\[
  \ell =
  - 0.5 α^{2} 
  - 0.5 β^{2}
  - σ^{2} 
  + \sum_{\left(j_{1}\right)=1}^{n} \left(- \log{\left(\sigma \right)} 
  - \frac{\log{\left(\pi \right)}}{2} 
  - \frac{\log{\left(2 \right)}}{2} 
  - \frac{\left({y}_{\left(j_{1}\right)} 
  - {\hat{y}}_{\left(j_{1}\right)}\right)^{2}}{2 \sigma^{2}}\right) 
  + \sum_{j_{1}=1}^{n} \left(- 0.005 {x}_{j_{1}}^{2} 
  - 2.99573227355399 
  - \frac{\log{\left(\pi \right)}}{2} 
  + \frac{\log{\left(2 \right)}}{2}\right) 
  - 2 \log{\left(\pi \right)} 
  - 1.38629436111989
\]

\[
  \ell =
  - 3.7 
  - 4.1 n 
  - \frac{ \alpha^{2} }{2}
  - \frac{\beta^2}{2}
  - \sigma^{2} 
  - n \log{\left(\sigma \right)} 
  - 0.005 \sum_{j_{1}=1}^{n} {x}_{j_{1}}^{2} 
  - \frac{1}{\sigma^2}\sum_{j_{1}=1}^{n} \left({y}_{j_{1}} 
  - {\hat{y}}_{j_{1}}\right)^{2}
\]

\subsection{Code Generation}
